---
title: "export_municipios_farmacias"
author: "Rodrigo Hernandez"
date: "2024-09-28"
output: html_document
---
```{r}
# Get rid of scientific notation
options(scipen = 99999999)

# Load the packages to be used
# First check is "pacman"is already installed and installs it otherwise
if (!require("pacman")) {
  install.packages("pacman")
}

#Installing and loading the packages needed
pacman::p_load(readr, dplyr, ggplot2, stringr, sf, RColorBrewer)
#Intalling janitor
install.packages("janitor")
library(janitor)


```

```{r}
farmacias_total_denue = read_csv("../../dataton2024_src/denue_00_46321-46531_csv/conjunto_de_datos/denue_inegi_46321-46531_.csv")|>
  mutate(nombre_act = str_to_lower(nombre_act))|>
  filter(str_detect(nombre_act, "farm"))
```
```{r}
#municipios 
Sys.setenv(OGR_GEOMETRY_ACCEPT_UNCLOSED_RING = "NO")
municipios = st_read("../../dataton2024_src/cartografia/MUNICIPIO.shp")
municipios <- st_transform(municipios, crs = 4326)

```

```{r}
frontera_iztapalapa = municipios |>
filter(ENTIDAD == 09, MUNICIPIO == 07)
print(filter(municipios, ENTIDAD == 09, MUNICIPIO == 07))
```

```{r}
ggplot()+
geom_sf(data = frontera_iztapalapa,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
iztapalapa_farmacias = farmacias_total_denue |>
filter(cve_ent == "09", cve_mun == "007")
print(filter(farmacias_total_denue,cve_ent == "09", cve_mun == "007"))
```

```{r}
iztapalapa_farmacias_sf <- st_as_sf(iztapalapa_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_iztapalapa = st_transform(iztapalapa_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_iztapalapa, fill = "white", color = "grey80")+
geom_sf(data = lambert_iztapalapa, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 
```{r}
filter(farm_municipios, entidad == 9, municipio == 7)
```
EXPORTACION
```{r}
st_write(lambert_iztapalapa, "../../exports/lambert_iztapalapa.shp")
#exportar frontera iztapalapa
st_write(frontera_iztapalapa, "../../exports/frontera_iztapalapa.shp")
```

#Ecatzingo

```{r}
frontera_ecatzingo = municipios |>
filter(ENTIDAD == 15, MUNICIPIO == 35)
print(filter(municipios, ENTIDAD == 15, MUNICIPIO == 35))
```

```{r}
ggplot()+
geom_sf(data = frontera_ecatzingo,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
print(filter(farmacias_total_denue, municipio == 'Ecatzingo'))

ecatzingo_farmacias = farmacias_total_denue |>
filter(cve_ent == "15", cve_mun == "034")
print(filter(farmacias_total_denue,cve_ent == "15", cve_mun == "034"))
```

```{r}
ecatzingo_farmacias_sf <- st_as_sf(ecatzingo_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_ecatzingo = st_transform(ecatzingo_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_ecatzingo, fill = "white", color = "grey80")+
geom_sf(data = lambert_ecatzingo, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 
```{r}
filter(farm_municipios, entidad == 15, municipio == 035)
```
EXPORTACION
```{r}
st_write(lambert_ecatzingo, "../../exports/lambert_ecatzingo.shp")
#exportar frontera iztapalapa
st_write(frontera_ecatzingo, "../../exports/frontera_ecatzingo.shp")
```
#Quecholac

```{r}
frontera_quecholac = municipios |>
filter(ENTIDAD == 21, MUNICIPIO == 116)
print(filter(municipios, ENTIDAD == 21, MUNICIPIO == 116))
```

```{r}
ggplot()+
geom_sf(data = frontera_quecholac,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
print(filter(farmacias_total_denue, municipio == 'Quecholac'))

quecholac_farmacias = farmacias_total_denue |>
filter(cve_ent == "21", cve_mun == "115")
print(filter(farmacias_total_denue,cve_ent == "21", cve_mun == "115"))
```

```{r}
quecholac_farmacias_sf <- st_as_sf(quecholac_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_quecholac = st_transform(quecholac_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_quecholac, fill = "white", color = "grey80")+
geom_sf(data = lambert_quecholac, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 
```{r}
filter(farm_municipios, entidad == 21, municipio == 116)
```
EXPORTACION
```{r}
st_write(lambert_quecholac, "../../exports/quecholac/lambert_quecholac.shp")
#exportar frontera iztapalapa
st_write(frontera_quecholac, "../../exports/quecholac/frontera_quecholac.shp")
```
#Nicolás Romero 

```{r}
frontera_nicolas_romero = municipios |>
filter(ENTIDAD == 15, MUNICIPIO == 61)
print(filter(municipios, ENTIDAD == 15, MUNICIPIO == 61))
```

```{r}
ggplot()+
geom_sf(data = frontera_nicolas_romero,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
print(filter(farmacias_total_denue, cve_ent == "15", cve_mun == "060"))

nicolas_romero_farmacias = farmacias_total_denue |>
filter(cve_ent == "15", cve_mun == "060")
print(nicolas_romero_farmacias)
```

```{r}

nicolas_romero_farmacias_sf <- st_as_sf(nicolas_romero_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_nicolas_romero = st_transform(nicolas_romero_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_nicolas_romero, fill = "white", color = "grey80")+
geom_sf(data = lambert_nicolas_romero, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 
```{r}
filter(farm_municipios, entidad == 15, municipio == 61)
```
EXPORTACION
```{r}
st_write(lambert_nicolas_romero, "../../exports/Nicolás Romero/lambert_nicolas_romero.shp")
#exportar frontera iztapalapa
st_write(frontera_nicolas_romero, "../../exports/Nicolás Romero/frontera_nicolas_romero.shp")
```
#Morelos  

```{r}
frontera_morelos = municipios |>
filter(ENTIDAD == 16, MUNICIPIO == 55)
print(filter(municipios, ENTIDAD == 16, MUNICIPIO == 55))
```

```{r}
ggplot()+
geom_sf(data = frontera_morelos,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
print(filter(farmacias_total_denue, cve_ent == "16", cve_mun == "054"))

morelos_farmacias = farmacias_total_denue |>
filter(cve_ent == "16", cve_mun == "054")
print(morelos_farmacias)
```

```{r}

morelos_farmacias_sf <- st_as_sf(morelos_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_morelos = st_transform(morelos_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_morelos, fill = "white", color = "grey80")+
geom_sf(data = lambert_morelos, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 
```{r}
filter(farm_municipios, entidad == 16, municipio == 55)
```
EXPORTACION
```{r}
st_write(lambert_morelos, "../../exports/Morelos/lambert_morelos.shp")
#exportar frontera iztapalapa
st_write(frontera_morelos, "../../exports/Morelos/frontera_morelos.shp")
```
#Morelos  

```{r}
frontera_morelos = municipios |>
filter(ENTIDAD == 16, MUNICIPIO == 55)
print(filter(municipios, ENTIDAD == 16, MUNICIPIO == 55))
```

```{r}
ggplot()+
geom_sf(data = frontera_morelos,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
print(filter(farmacias_total_denue, cve_ent == "16", cve_mun == "054"))

morelos_farmacias = farmacias_total_denue |>
filter(cve_ent == "16", cve_mun == "054")
print(morelos_farmacias)
```

```{r}

morelos_farmacias_sf <- st_as_sf(morelos_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_morelos = st_transform(morelos_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_morelos, fill = "white", color = "grey80")+
geom_sf(data = lambert_morelos, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 
```{r}
filter(farm_municipios, entidad == 16, municipio == 55)
```
EXPORTACION
```{r}
st_write(lambert_morelos, "../../exports/Morelos/lambert_morelos.shp")
#exportar frontera iztapalapa
st_write(frontera_morelos, "../../exports/Morelos/frontera_morelos.shp")
```

#Donato Guerra, Tlataya y Tonatico
municipality table for remaining municipalities from Donato Guerra to Leon 

First manual job: find the municipalities in "municipios" and in denue. Fill in their entidad, municipio and cve_ent, cve_mun respectively. 
```{r}
municipality_table_1 <- data.frame(
  ENTIDAD = c(15, 15, 15),
  MUNICIPIO = c(33, 108, 106),
  cve_ent = c("15", "15", "15"),
  cve_mun = c("032", "107", "105"),
  name = c("Donato Guerra", "Tonatico", "Tlatlaya")
)
```

```{r}
print(filter(farmacias_total_denue, cve_ent == "19", cve_mun == "039"))
```

```{r}
# Loop through the table and run the code for each municipality
for (i in 1:nrow(municipality_table_1)) {
  
  # Extract data for the current municipality
  entidad <- municipality_table_1$ENTIDAD[i]
  municipio <- municipality_table_1$MUNICIPIO[i]
  ent_code <- municipality_table_1$cve_ent[i]
  mun_code <- municipality_table_1$cve_mun[i]
  name <- municipality_table_1$name[i]
  
  # Filter for the selected municipality
  frontera_municipio <- municipios |>
    filter(ENTIDAD == entidad, MUNICIPIO == municipio)
  
  # Plot the municipality boundary
  ggplot() +
    geom_sf(data = frontera_municipio, color = "grey", fill = "white", size = 0.1) +
    theme_minimal()
  
  # Filter drug stores in the municipality
  print(filter(farmacias_total_denue, cve_ent == cve_ent, cve_mun == mun_code))
  
  municipio_farmacias <- farmacias_total_denue |>
    filter(cve_ent == ent_code, cve_mun == mun_code)
  print(municipio_farmacias)
  
  # Convert to spatial data
  municipio_farmacias_sf <- st_as_sf(municipio_farmacias, coords = c("longitud", "latitud"), crs = 4326)
  
  # Transform coordinates
  lambert_municipio <- st_transform(municipio_farmacias_sf, crs = 4326)
  
  # Plot the municipality with drug store locations
  ggplot() +
    geom_sf(data = frontera_municipio, fill = "white", color = "grey80") +
    geom_sf(data = lambert_municipio, color = "blue", alpha = 1, size = 1) +
    theme_minimal()
  
  # Test number of drug stores by municipality
  filter(farm_municipios, entidad == entidad, municipio == municipio)
  
  # Export the spatial data
  st_write(lambert_municipio, paste0("../../exports/", name, "/lambert_", name, ".shp"))
  st_write(frontera_municipio, paste0("../../exports/", name, "/frontera_", name, ".shp"))
  
}
```
```{r}
  ggplot() +
    geom_sf(data = lambert_donato, fill = "white", color = "grey80") +
    geom_sf(data = lambert_municipio, color = "blue", alpha = 1, size = 1) +
    theme_minimal()
```
#Ixtapan de la Sal   

```{r}
frontera_ixtapan = municipios |>
filter(ENTIDAD == 15, MUNICIPIO == 41)
print(filter(municipios, ENTIDAD == 15, MUNICIPIO == 41))
```

```{r}
ggplot()+
geom_sf(data = frontera_morelos,
color = "grey",
fill = "white",
size = 0.1)+
theme_minimal()
```

```{r}
print(filter(farmacias_total_denue, cve_ent == "15", cve_mun == "040"))

morelos_farmacias = farmacias_total_denue |>
filter(cve_ent == "15", cve_mun == "040")
print(morelos_farmacias)
```

```{r}

morelos_farmacias_sf <- st_as_sf(morelos_farmacias, coords = c("longitud", "latitud"), crs = 4326)

lambert_morelos = st_transform(morelos_farmacias_sf, crs = 4326)

```

```{r}
ggplot()+
geom_sf(data = frontera_morelos, fill = "white", color = "grey80")+
geom_sf(data = lambert_morelos, color = "blue", alpha = 1, size = 1)+
theme_minimal()
```

Prueba cuenta de farm por muni: 


```{r}
municipality_table_2 <- data.frame(
  ENTIDAD = c(15, 15, 15, 15, 26, 7, 9, 23, 11, 13),
  MUNICIPIO = c(41, 111, 83, 62, 67, 101, 5, 2, 20, 46),
  cve_ent = c("15", "15", "15", "15", "26", "07", "09", "23", "11", "13"),
  cve_mun = c("040", "110", "082", "061", "049", "102", "005", "001", "020", "047"),
  name = c("ixtapan_de_la_sal", "Valle de Bravo", "Tejupilco", "Nopaltepec", "Quiriego", "Tuxla_chico", "gustavo_madero", "Cozumel", "Leon", "pacula")
)
```

```{r}
print(filter(farmacias_total_denue, cve_ent == "13", cve_mun == "047"))
```

```{r}
# Loop through the table and run the code for each municipality
for (i in 1:nrow(municipality_table_2)) {
  
  # Extract data for the current municipality
  entidad <- municipality_table_2$ENTIDAD[i]
  municipio <- municipality_table_2$MUNICIPIO[i]
  ent_code <- municipality_table_2$cve_ent[i]
  mun_code <- municipality_table_2$cve_mun[i]
  name <- municipality_table_2$name[i]
  
  # Filter for the selected municipality
  frontera_municipio <- municipios |>
    filter(ENTIDAD == entidad, MUNICIPIO == municipio)
  
  # Plot the municipality boundary
  ggplot() +
    geom_sf(data = frontera_municipio, color = "grey", fill = "white", size = 0.1) +
    theme_minimal()
  
  # Filter drug stores in the municipality
  print(filter(farmacias_total_denue, cve_ent == cve_ent, cve_mun == mun_code))
  
  municipio_farmacias <- farmacias_total_denue |>
    filter(cve_ent == ent_code, cve_mun == mun_code)
  print(municipio_farmacias)
  
  # Convert to spatial data
  municipio_farmacias_sf <- st_as_sf(municipio_farmacias, coords = c("longitud", "latitud"), crs = 4326)
  
  # Transform coordinates
  lambert_municipio <- st_transform(municipio_farmacias_sf, crs = 4326)
  
  # Plot the municipality with drug store locations
  ggplot() +
    geom_sf(data = frontera_municipio, fill = "white", color = "grey80") +
    geom_sf(data = lambert_municipio, color = "blue", alpha = 1, size = 1) +
    theme_minimal()
  
  # Test number of drug stores by municipality
  filter(farm_municipios, entidad == entidad, municipio == municipio)
  
  # Export the spatial data
  st_write(lambert_municipio, paste0("../../exports/", name, "/lambert_", name, ".shp"))
  st_write(frontera_municipio, paste0("../../exports/", name, "/frontera_", name, ".shp"))
  
}
```
```

